\bf Accelerated and interpretable oblique random survival forests

### 2023-05-25 paper/jcgs/main.tex
R2 and AE updates

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/axis_v_oblique-1} \caption[Decision boundaries from an axis based (panel A) and oblique (panel B) decision tree used to classify penguin species based on bill depth and bill length]{Decision boundaries from an axis based (panel A) and oblique (panel B) decision tree used to classify penguin species based on bill depth and bill length. The decision boundary from the oblique tree is better able to capture the geometry of this data, leading to fewer mis-classified penguins.}\label{fig:axis_v_oblique-1}
\end{figure}

\begin{figure}
\includegraphics[width=\maxwidth]{figure/axis_v_oblique-2} \caption[Decision boundaries from an axis based (panel A) and oblique (panel B) decision tree used to classify penguin species based on bill depth and bill length]{Decision boundaries from an axis based (panel A) and oblique (panel B) decision tree used to classify penguin species based on bill depth and bill length. The decision boundary from the oblique tree is better able to capture the geometry of this data, leading to fewer mis-classified penguins.}\label{fig:axis_v_oblique-2}
\end{figure}

@@

 \item The ``VA lung cancer trial'' data \citep{kalbfleisch2011statistical} were obtained from the \texttt{randomForestSRC} R package \citep{randomForestSRC}. 
 \item The ``Colon cancer'' data \citep{moertel1995fluorouracil} were obtained from the \texttt{survival} R package \citep{survival}. 
 \item The ``Primary biliary cholangitis'' data \citep{therneau2000cox} were obtained from the \texttt{aorsf} R package \citep{aorsf}. 
 \item The ``Movies released in 2015-2018'' data  were obtained from the \texttt{censored} R package \citep{censored}. 
 \item The ``GBSG II'' data \citep{schumacher1994rauschecker} were obtained from the \texttt{TH.data} R package \citep{TH.data}. 
 \item The ``Systolic Heart Failure'' data \citep{hsich2011identifying} were obtained from the \texttt{randomForestSRC} R package \citep{randomForestSRC}. 
 \item The ``Serum free light chain'' data \citep{dispenzieri2012use, kyle2006prevalence} were obtained from the \texttt{survival} R package \citep{survival}. 
 \item The ``Non-alcohol fatty liver disease'' data \citep{allen2018nonalcoholic} were obtained from the \texttt{survival} R package \citep{survival}. 
 \item The ``Rotterdam tumor bank'' data \citep{royston2013external} were obtained from the \texttt{survival} R package \citep{survival}. 
 \item The ``ACTG 320'' data \citep{hosmer2002applied} were obtained from the \texttt{mlr3proba} R package \citep{mlr3proba}. 
 \item The ``GUIDE-IT'' data \citep{felker2017effect} were obtained from BioLINCC. 
 \item The ``Early breast cancer'' data \citep{desmedt2011multifactorial, hatzis2011genomic, ternes2017identification} were obtained from the \texttt{biospear} R package \citep{biospear}. 
 \item The ``SPRINT'' data \citep{sprint2015randomized} were obtained from BioLINCC. 
 \item The ``NKI 70 gene signature'' data \citep{van2002gene} were obtained from the \texttt{OpenML} R package \citep{OpenML}. 
 \item The ``Lung cancer'' data \citep{director2008gene} were obtained from the \texttt{OpenML} R package \citep{OpenML}. 
 \item The ``NCCTG Lung Cancer'' data \citep{loprinzi1994prospective} were obtained from the \texttt{survival} R package \citep{survival}. 
 \item The ``FCL'' data \citep{pintilie2006competing} were obtained from the \texttt{randomForestSRC} R package \citep{randomForestSRC}. 
 \item The ``Monoclonal gammopathy'' data \citep{kyle2002long} were obtained from the \texttt{survival} R package \citep{survival}. 
 \item The ``MESA'' data \citep{bild2002multi} were obtained from BioLINCC. 
 \item The ``ARIC'' data \citep{aric1989atherosclerosis} were obtained from BioLINCC. 
 \item The ``JHS'' data \citep{taylor2005toward} were obtained from BioLINCC.

>>>>

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/axis_v_oblique-1} \caption[Decision boundaries from an axis based (panel A) and oblique (panel B) decision tree used to classify penguin species based on bill depth and bill length]{Decision boundaries from an axis based (panel A) and oblique (panel B) decision tree used to classify penguin species based on bill depth and bill length. The decision boundary from the oblique tree is better able to capture the geometry of this data, leading to fewer mis-classified penguins.}\label{fig:axis_v_oblique}
\end{figure}

@@

 \item The ``VA lung cancer trial'' data \citep{kalbfleisch2011statistical} were obtained from the \texttt{randomForestSRC} R package \citep{randomForestSRC}. No relevant pre-processing steps were taken. 
 \item The ``Colon cancer'' data \citep{moertel1995fluorouracil} were obtained from the \texttt{survival} R package \citep{survival}. In pre-processing steps, the predictor, node4, which is an indicator for having more than 4 positive lymph nodes, was dropped, while the nodes predictor, which indicates the number of positive lymph nodes, was retained. 
 \item The ``Primary biliary cholangitis'' data \citep{therneau2000cox} were obtained from the \texttt{aorsf} R package \citep{aorsf}. No relevant pre-processing steps were taken. 
 \item The ``Movies released in 2015-2018'' data  were obtained from the \texttt{censored} R package \citep{censored}. In pre-processing steps, text and date variables (movie title, release date, and distributor) were dropped.. 
 \item The ``GBSG II'' data \citep{schumacher1994rauschecker} were obtained from the \texttt{TH.data} R package \citep{TH.data}. No relevant pre-processing steps were taken. 
 \item The ``Systolic Heart Failure'' data \citep{hsich2011identifying} were obtained from the \texttt{randomForestSRC} R package \citep{randomForestSRC}. No relevant pre-processing steps were taken. 
 \item The ``Serum free light chain'' data \citep{dispenzieri2012use, kyle2006prevalence} were obtained from the \texttt{survival} R package \citep{survival}. In pre-processing steps, the chapter variable, which indicates death status, was removed, since death was the outcome. Outcomes occurring on day 0 were assumed to have a time of 1/2 day rather than 0 days. 
 \item The ``Non-alcohol fatty liver disease'' data \citep{allen2018nonalcoholic} were obtained from the \texttt{survival} R package \citep{survival}. In pre-processing steps, data from before or on the index data were used as predictors. The mean value prior to the index date for lab values in nafld2 was used as a predictor, and the number of days between the most recent lab test and the index date was also used as a predictor. 
 \item The ``Rotterdam tumor bank'' data \citep{royston2013external} were obtained from the \texttt{survival} R package \citep{survival}. No relevant pre-processing steps were taken. 
 \item The ``ACTG 320'' data \citep{hosmer2002applied} were obtained from the \texttt{mlr3proba} R package \citep{mlr3proba}. In pre-processing steps, redundant predictors were dropped. 
 \item The ``GUIDE-IT'' data \citep{felker2017effect} were obtained from BioLINCC. In pre-processing steps, baseline data including biomarkers, questionnaires, and randomized group were included as predictors. 
 \item The ``Early breast cancer'' data \citep{desmedt2011multifactorial, hatzis2011genomic, ternes2017identification} were obtained from the \texttt{biospear} R package \citep{biospear}. No relevant pre-processing steps were taken. 
 \item The ``SPRINT'' data \citep{sprint2015randomized} were obtained from BioLINCC. In pre-processing steps, baseline data including biomarkers, cognitive questionnaires, medications, and randomization group were included as predictors. Predictors with over 40 percent missing data were dropped. Zero variance predictors were also dropped. 
 \item The ``NKI 70 gene signature'' data \citep{van2002gene} were obtained from the \texttt{OpenML} R package \citep{OpenML}. No relevant pre-processing steps were taken. 
 \item The ``Lung cancer'' data \citep{director2008gene} were obtained from the \texttt{OpenML} R package \citep{OpenML}. In pre-processing steps, status was transformed to have values of 0 and 1 instead of 1 and 2. 
 \item The ``NCCTG Lung Cancer'' data \citep{loprinzi1994prospective} were obtained from the \texttt{survival} R package \citep{survival}. In pre-processing steps, institution code was not used as a predictor, and values of both sex and event status were transformed to be 0 and 1 instead of 1 and 2. 
 \item The ``FCL'' data \citep{pintilie2006competing} were obtained from the \texttt{randomForestSRC} R package \citep{randomForestSRC}. No relevant pre-processing steps were taken. 
 \item The ``Monoclonal gammopathy'' data \citep{kyle2002long} were obtained from the \texttt{survival} R package \citep{survival}. No relevant pre-processing steps were taken. 
 \item The ``MESA'' data \citep{bild2002multi} were obtained from BioLINCC. In pre-processing steps, visit 1 data including biomarkers, health behaviors, and comorbidities were included as predictors. 
 \item The ``ARIC'' data \citep{aric1989atherosclerosis} were obtained from BioLINCC. In pre-processing steps, visit 1 data including biomarkers, health behaviors, and comorbidities were included as predictors. 
 \item The ``JHS'' data \citep{taylor2005toward} were obtained from BioLINCC. In pre-processing steps, visit 1 data including biomarkers, health behaviors, neighborhood characteristics, and comorbidities were included as predictors.
